MLH1 Data Report

### 2019-12-17 doc/Main_Report.rmd
missed some updating

#load main data file
load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

@@

## Discription of quantification process

@@

```{r organizing.data, echo=FALSE}
#set the order of another column, based on another variable (so that when)

@@

This section shows the number of mice which haven't been stained or quantified yet.

```{r missing.mice, echo=FALSE, warning=FALSE}

@@

passed quantified - quantified mice (with enough cells)
total quantified -- all mice which have been (includes fail and passing mice)

non-quantified mice -- dissected mice / slides that haven't been quantified yet (all stained, but without image directory)

non-imaged -- mice without images, (might be stained and unstained)

This is the number of mice that are through different parts of the pathway or quantification process.

should I break down the categories into sex and strain?

@@

These are the tables for the number of cells and mice which have been quantified.
The 'Other' are F1 crosses.

## Repeatability of Human quantification

Well, we have to talk about the subjective nature that goes into scoring the MLH1 foci. Sources of error or variance across scoring come from background noise (false positives), faint foci (false negatives) and just human error.
A general anecdote is that if the quality of the cell staining / image is high, then the repeatability is near 100%. (using some of Beth's data).
Todo: track down, re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)

@@

Datasets / sources of repeat measures
-chrom proportions
-multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
-old and new (batch0)

The main take away is that higher quality scored cells have lower error in MLH1 counts.

@@

```{r quant.repeatability_rep1, echo=FALSE, include=TRUE}
#set up data for combining replicate counts

#1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now

@@

levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt

#check that the 1, and 2 are equal number
length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178

@@

#i don't know how to compare quality in more

@@

```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE}
#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
#load, clean up and mat one of the master.MLH1count sheets
#master
old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
#remove all with 'X'
#create a real fileName col, add rev.tif, 
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

str(test.old.master)

@@

```{r repeatability_rep3, echo=FALSE, include=FALSE}
chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#what are the target columns for this excel sheet?
#espically for getting repeat data
#MLH1.dif  same_count  quality

@@

This data has `r length(chrm.class.DF$Original.Name)` cells across these categories

@@

More cells are planned to be quantified.

```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE}
hist_error_plot
```

From this histogram we see most of the cells in this dataset have a difference of 2 in the MLH1 count scores (`r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`).

```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5}
quality_error_plot
```

There is a positive relationship with the absolute difference and quality score (where 1 is best and 5 is worst quality) (i.e. poorer quality cells are more likely to have more error acorss multiple counts). The proportion of the cells which had the same count across the counting instances is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`. The correlation between the two versions of MLH1 counts for the same cell is `r cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)`.

```{r show.error_count_plot, echo=FALSE,  fig.width=10, fig.height=5}
error_count_plot
```
There doesn't seem to be a bias across cells with more or less foci.

Modeling this relationship; (still in progress)
I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.

## Effects on the dataset

error associated with 'quality of cels'

Main metric is mean CO count per mouse. Run tests to check if there are effects for the following effects.

*age effects* (male, maternal age)

*litter effect*

*mouse room effect* (Charmony, MSC)

*quantification repeatability*

*spread batch and staining effects*

*power analysis on number of cells*

*time of year effect (on spread)*

*dominant effect* (F1s)

These aspects will be analyzed with a linear models on subsets of data (within category) and visualized with plots.

@@

There is a positive correlation with MLH1 number and quality
To test this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.

@@

Q12toTotal

@@

kable(Q12_MLH1_means_males, caption = "High quality cells")

kable(tot.DF_MLH1_means_males, caption = "Total Cell")

@@

(try looking at <5 quality scores?)

@@

### Age Effect

@@

Make a subsetted dataset for plotting mean mouse (quality < 5) MLH1 count and age.

@@

I will try doing a logistic regression for embryonic vs neonate. The embyonic estimates likely aren't very accurate, and there won't be good sampling across time points.

@@

@@

#check if date format's been applied before tryung the age functions
str(meta.data.file.w.mat.DOB)

@@

str(meta.data.file.w.mat.DOB)

@@

#merge meta data with MLH1 
see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
#remove anything without mean_CO
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
#what did I do with this DF...?

#visualize the data

meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]

@@

#age by mean_co
mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
mat.age.effect.plot.female

@@

There seems to be a negative correlation with

@@

Compare female spreads made on the same day, but from different mothers/litters.

@@

### Room effect (envioronment)

3 different rooms used to house mice. Access softmouse to be fixed. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD

@@

# Initial Results (change these colors)

@@

#the below ordering works!! keep it
MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
              "PWD","MSM","MOLF","CZECH","SKIVE",
              "KAZ","TOM","AST",
              "CAST","HMI","SPRET","SPIC", "other"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

#don't know why this won't do the right order...
#set the order of another column, based on another variable (so that when)

#try to make sure sacles are the same

#reduncant
#MLH1_data <- MLH1_data %>%
 # arrange(strain, sex, mouse) %>%
#  mutate(Original.Name = factor(Original.Name)) #another category that you want the order to match
# sort your dataframe, by the focal categories
#MLH1_data$mouse <- factor(MLH1_data$mouse, levels=unique(MLH1_data$mouse))

#try this for ordering
#MLH1_data[order(MLH1_data$strain), ]

#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
#can't re-order the miceagg
#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
#        "grey","grey","grey",  "grey","grey","grey" )) +
# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
#don't show M.musculus by mouse

#order MLH1_data by strain!!
#colors_of_strains, colors ..

ff2 <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
  geom_boxplot(aes(x = as.character(mouse), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
        "green","green4","grey",  "grey","grey","grey" )) +
 ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) + 
  facet_grid(subsp~ sex, scales="free_x")+scale_color_manual(values=colors_of_strains)

@@

@@

```{r MLH1.histogram, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}

(most all distributions are close to normal)

How do these compare to Poisson? (mean == variance)
I should make 'expected' distributions behind these.

```{r passing.mice, echo=FALSE, warning=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table
passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)

```{r scatter.plots.of.nMLH1.by.score, echo=FALSE, fig.cap="MLH1 means and quality scores", warning=FALSE}

#make lattice / plots seperated by strain and sex
#facet

#quality scores are now 1:7?
qual_means <- MLH1_data %>% 
        group_by(category, quality) %>% 
        summarise(
          adj_nMLH1.foci = mean(adj_nMLH1.foci)
        )

p <- ggplot(MLH1_data,(aes(adj_nMLH1.foci,quality)))
p <- p + geom_point(stat = "identity")
p <- p  + geom_point(data = qual_means, colour="red") 
p <- p + facet_wrap(~ category)
p

####Distributions by mouse

@@

### Explaination for the XX and XY differences

(in progress)

@@

generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
10 cells from each musc category.

@@

### Mixed Model, compare within and between subspecies patterns

@@

  variables vs char coding of coefficents(opposite direction of sex)?
  select1 (remove old mice, remove mice with low ncells)
  select2 make sure there are equal numbers of observations
  select3 low Musc males

the pooled cell models are breaking for some reason, (the sexes)
-pooled cells (nMLH1)
  all 
  pooled.HQ (1-4)
  pooled.HQ.2 (sample equal numbers of cells)

```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE}
library(lme4)
library(nlme)
# seperate the data, both sexes, only Mus
# decide on Musc - Molossinus
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

@@

#if(grepl("MSM", DF.HetC.MixedModel2$strain) ){ DF.HetC.MixedModel2$subsp <- "Mol" }

#subsp are charecters
#change sex to charecter? (unordered sex)

DF.HetC.MixedModel2$sex <- as.character(DF.HetC.MixedModel2$sex)

model.ff2 <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel2, random=list(strain=pdDiag(~sex) ) )#adding mol

@@

#intercept should be the Dom female mean, looks right
mean(DF.HetC.MixedModel2$mean_co[(DF.HetC.MixedModel2$sex == "female" &
                                    DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5

@@

#remove PERC, MOLF, CZECH,
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "CZECH",]
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "MOLF",]
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "PERC",]
```

@@

```{r mixed.model.make, echo=FALSE, warning=FALSE, eval=FALSE, message=FALSE}
#build mixed model
#run simulations for accounting for sampling error
# old code uses base plot
library(nlme)
library(lme4)

#L,Q,C  appear for ordered factors in the summary mean lme doesn't believe you used ordered factors
#change all the DFs / variables to UNORDERED factors!
#redo the orders just to be sure
#sex, strain, subsp,

DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE, levels =c("female", "male") )

DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE, levels =c( "Dom", "Musc") )

DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW",
                                                  "PWD", "MSM", "SKIVE", "KAZ") )

>>>>

#load main data file

load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

@@

@@

```{r organizing.data, echo=FALSE }
#set the order of another column, based on another variable (so that when)

@@

```{r missing.mice, echo=FALSE, warning=FALSE, eval=FALSE}
#This section shows the number of mice which haven't been stained or quantified yet.

@@

@@

```{r rand, echo=FALSE, include=FALSE}

str(AP_mouse_table_w.Ages)

AP_mouse_table.HQ$var <- as.numeric(AP_mouse_table.HQ$var)

#remove sex.x == na, and strain.x=na or F1
AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$sex) | AP_mouse_table.HQ$sex==""), ]

AP_mouse_table.HQ <- AP_mouse_table.HQ[!(is.na(AP_mouse_table.HQ$var) | AP_mouse_table.HQ$var==""), ]
#plot variance and mean (facet sex)

#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!( AP_mouse_table_w.Ages$strain.x =="F1"), ]

VandMean.mouse <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; variance by mean")

cVandMean.mouse <- ggplot(aes(x=mean_co, y = cV), data=AP_mouse_table.HQ )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("mouse MLH1 averages; cV by means")

str(AP_mouse_table.HQ)

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "female"])
#[1] 0.5750399
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"])
#0.321601

cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "female"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "female"])
# 0.3515544
cor(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"])
#-0.03713536

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$var[AP_mouse_table.HQ$sex == "male"] )
# p-value = 8.623e-09

cor.test(AP_mouse_table.HQ$mean_co[AP_mouse_table.HQ$sex == "male"], 
    AP_mouse_table.HQ$cV[AP_mouse_table.HQ$sex == "male"] )

#remove mice with fewer than 10 cells

AP_mouse_table.HQ.10plus <- AP_mouse_table.HQ[AP_mouse_table.HQ$Ncells >= 10,]

VandMean.mouse.10plus <- ggplot(aes(x=mean_co, y = var), data=AP_mouse_table.HQ.10plus )+geom_point(aes(color=strain))+facet_wrap(~sex)+ggtitle("10 plus mouse MLH1 averages; variance by mean")

cor(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], AP_mouse_table.HQ.10plus$cV[AP_mouse_table.HQ.10plus$sex == "female"])

cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )

## Repeatability of Human Quantification

Let's talk about the subjective nature of scoring the MLH1 foci. Sources of error or variance across scoring come from background noise (false positives), faint foci (false negatives) and just human error. A general anecdote is that if the quality of the cell staining is high, then the repeatability is near 100%. When the quality is less than perfect, some foci can ambigous and cause the repeatability to decrease.

@@

-re-count of Beth's data,  cells that were counted first then in batch 0 (anonimized)
Datasets / sources of repeat measures
-cell counts from chrm proportions vs batch
-multiple batch counts (later than batch0) *focus on this dataset for now. since its >100 and fairly easy to compile
-old and new (old counts and anonimized batch0)

Some points I still need to address

@@

```{r quant.repeatability_rep1, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE}
#set up data for combining replicate counts
#1. replicates in DF, (17 mice, across 10 batches), focus on those from original for now

@@

#levels(matches.only$nMLH1.foci) #leves don't appear in num?.. move on, dwabt

#check that the 1, and 2 are equal number
#length(matches.only$Original.Name[matches.only$separate.DF == 1]) #both 178

@@

@@

```{r repeatability_rep2, echo=FALSE, include=FALSE, eval=FALSE, message=FALSE}
#2. old vs new, (remove duplicates and merge batch0 with old_master_mlh1)
#load, clean up and mat one of the master.MLH1count sheets
#master
old.master.MLH1 = read.csv("~./MLH1repo/data/old_master_MLH1_nov15.csv", header = TRUE)
#remove all with 'X'
#create a real fileName col, add rev.tif, 
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

#str(test.old.master)

@@

```{r repeatability_rep3, echo=FALSE, include=FALSE, message=FALSE}
chrm.class.DF = read.csv("~./MLH1repo/data/ChrmClassProportions.csv", header = TRUE)
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#what are the target columns for this excel sheet?
#espically for getting repeat data
#MLH1.dif  same_count  quality

@@

This data has `r length(chrm.class.DF$Original.Name)` cells from mice across these categories

@@

```{r show.hist_error_plot, echo=FALSE,  fig.width=10, fig.height=5, warning=FALSE, message=FALSE}
hist_error_plot
```

From this histogram we see most of the cells in this dataset have a difference of less than 2, or `r length(chrm.class.DF$Original.Name[( abs( chrm.class.DF$MLH1.dif) <= 2)]) / length(chrm.class.DF$Original.Name)`%.

```{r show.quality_error_plot, echo=FALSE,  fig.width=10, fig.height=5, message=FALSE}
quality_error_plot
```

There is a positive relationship with the absolute difference and quality score (where 1 is best and 5 is worst quality) (i.e. poorer quality cells are more likely to have more error acorss multiple counts). The proportion of the cells which had the same count is `r length(chrm.class.DF$Original.Name[chrm.class.DF$same_count == "TRUE"]) / length(chrm.class.DF$Original.Name)`%. The correlation between the two versions of MLH1 counts for the same cell is `r cor(chrm.class.DF$nMLH1.foci, chrm.class.DF$MLH1.sum.check)`.

```{r show.error_count_plot, echo=FALSE,  fig.width=10, fig.height=5}
error_count_plot
```
There doesn't seem to be a bias across cells with more foci.

```{r model.repeat, include=FALSE, eval=FALSE}
#add model -- logisitc regression?
#I tried doing a logistic regression for same_count = true or false depending on the quality score. I am having trouble making a plot the makes sense to confirm the binomial nature of the model.
```

```{r Effects.on.the.dataset, eval=FALSE, echo=FALSE}

#*age effects* (male, maternal age)

#*litter effect*

#*mouse room effect* (Charmony, MSC)

#*quantification repeatability*

#*spread batch and staining effects*

#*power analysis on number of cells*

#*time of year effect (on spread)*

#*dominant effect* (F1s)

@@

There is a positive correlation with MLH1 number and quality. To quantify this effect, compare all of the counts from quality-1 cells across strains. check that the same patterns of means persist compared to the total-cell data.

@@

#Q12toTotal

@@

#kable(Q12_MLH1_means_males, caption = "High quality cells")

#kable(tot.DF_MLH1_means_males, caption = "Total Cell")

@@

@@

### Age Effect (male)

@@

@@

@@

### Age Effect (female)

Compare embryonic to neonate (logistic regression). I don't think my embyonic stage estimates are very accurate, so I'll bin all embryos into 1.

```{r female.age, eval=FALSE}
#compile this dataset
```

@@

#check if date format's been applied before tryung the age functions
#str(meta.data.file.w.mat.DOB)

@@

#str(meta.data.file.w.mat.DOB)

@@

#merge meta data with MLH1 -- rename this
see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
#remove anything without mean_CO
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]
#what did I do with this DF...?

#visualize the data
meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]

@@

#age by mean_co (these aren't working)
mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
mat.age.effect.plot.female

@@

@@

Compare female spreads made on the same day, but from different mothers/litters.
```{r litter.effect, eval=FALSE}
#compile this data set

@@

```{r stain.effects, eval=FALSE}

#compile this dataset

### Room effect

```{r ROOM.effect, eval=FALSE}

#COMPILE THIS datasets
#3 different rooms used to house mice. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD

@@

```{r MLH1.histogram.2421, echo=FALSE, fig.caption= "histograms", warning=FALSE, fig.align='center', eval=FALSE}

```{r passing.mice.dont.use, echo=FALSE, warning=FALSE}
q_cutoff_table <- ddply(MLH1_data, .(mouse), summarise,
                        total =  length(adj_nMLH1.foci),
                        q5 = sum(as.numeric(quality) >= 4, na.rm = TRUE ), 
                        above5 = sum(as.numeric(quality) <= 4, na.rm = TRUE )
                       # q_l3 = sum(as.numeric(as.numeric(quality)) <= 4, na.rm = TRUE )
)
#q_cutoff_table
passed_mice <- as.vector(q_cutoff_table[q_cutoff_table$above5 > 15,]$mouse)

# Initial Results

@@

#the below ordering works!! keep it
MLH1_data$strain<- factor(MLH1_data$strain, levels =c("G", "LEW","WSB",
              "PWD","CZECH","SKIVE",
              "KAZ","TOM","AST",
              "MSM","MOLF",
              "CAST","HMI","SPRET","SPIC", "other"), order=T )
#order the data frame
MLH1_data <- with(MLH1_data, MLH1_data[order(sex, strain),])

#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
#can't re-order the miceagg
#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
#        "grey","grey","grey",  "grey","grey","grey" )) +
# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
#don't show M.musculus by mouse
#scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
 #       "green","green4","grey",  "grey","grey","grey" )) +

#order MLH1_data by strain!!
#colors_of_strains, colors .. 
#there is no order to 'mouse' so it looks shitty

justMuscu <- ggplot(MLH1_data[MLH1_data$species == "M.musculus", ])+
  geom_boxplot(aes(x = as.factor(strain), y = adj_nMLH1.foci, fill= strain,alpha=0.5) )+ ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
 ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
#facet_grid(~sex, scales="free_x")+scale_color_manual(values=colors_of_strains)+

justMuscu <- justMuscu   + facet_wrap(~ sex, scales="free")

@@

```{r variance.boxplots, echo=FALSE}
#plot of variance

var.boxplot <- ggplot(DF.HetC.MixedModel.HQ, aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Boxplots of MLH1 distributions") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
 ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+3)) +   scale_fill_manual(values=colors_of_strains)
var.boxplot <- var.boxplot   + facet_wrap(sex ~ strain, scales="free")

#plot by sex / seperately
# HQ data DF.HetC.MixedModel.HQ

var.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = var, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )

var.male <- var.male   + facet_wrap( ~ strain, scales="free")

cV.male <- ggplot(DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$sex == "male",], aes(x = as.factor(mouse), y = cV, fill= strain,alpha=0.5))+ geom_point()+
 ggtitle("Male variance levels") + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_fill_manual(values=colors_of_strains)+  ylim( c( 5, 20) )

cV.male <- cV.male   + facet_wrap( ~ strain, scales="free")

There are 2 MSM males with high (20) variance and cV, G seems to have more variance across mouse level variance too.

Check out these mice in comparison to other mice
30sep16_MSM_m1 -- staining might not be the best, MLH1 signal is spottiny in some cells

30may17_MSM_m1 - quite a bit of green noise across these cells.

Gough and KAZ males has a bit more variation across mice for mouse level variation.

@@

#### Distributions by mouse

@@

@@

generate smaller data set of chrm proportions (0CO, 1CO,2CO ect)
10 cells from each musc category. Add Molf male.

@@

todo: make model with strain as fixed effect (since, there are interesting strain effects)

@@

```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE, message=FALSE}
library(lme4)
library(nlme)

@@

@@

@@

@@

1. mean_co (chr coded)
2. mean_co.HQ
3. mean_co with Mol
4. sub-sampling

```{r mixed.models.setup, echo=FALSE}
library(raster)
#make are the DFs for the models
# seperate the data, both sexes, only Mus
# decide on Musc - Molossinus
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.x, sex.y,
                                                            subsp.y,subsp.x,
                                                            category.y, category.x,
                                                            strain.x,strain.y, mouse.1  )  )

#make dummy MOLF data, average of Musc females?
#no MOLF male data?
female.musc.cells <- MLH1_data[( MLH1_data$sex == "female") & (MLH1_data$subsp == "Musc" ), ]
table(female.musc.cells$quality)

female.musc.mice.table <- ddply(female.musc.cells, c("sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = format(round(   var(adj_nMLH1.foci),3), nsmall=3),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)#just 1 entry

female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
female.musc.mice.table$DOB <- as.Date("1/1/2020")

female.musc.mice.table$strain <- "MOLF"
female.musc.mice.table$subsp <- "Musc"
female.musc.mice.table$species <- "M.musculus"

female.musc.mice.table$female.stage <- "emb"
female.musc.mice.table$age.days <-""
female.musc.mice.table$age.weeks <- ""
female.musc.mice.table$ID <- ""
female.musc.mice.table$notes <- ""
female.musc.mice.table$euth.date <- ""

DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)

DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]

#main df with variables coded as chr
DF.HetC.MixedModel <- DF.HetC.MixedModel
DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE)
DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE)
DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE)

#mean_co.HQ  (this might remove some of the MOLF males)
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))

DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]

#sub-sample (need to think about this one)
